---
title: |
  | Outdoor Water Efficiency Survey:
  | Household Sample Creation
output: 
  html_document: default
  github_document: default
---

## Sample Size Calculation

Using the sample size estimation formula provided in a [UN handbook on household survey samples](http://unstats.un.org/unsd/demographic/sources/surveys/Handbook23June05.pdf) we have the following for the sample size $n_h$

$$n_h = (z^2)(r)(1-r)(f)(k)/(p)(\hat{n})(e^2)$$

Where

* $n_h$ is the parameter to be calculated and is the sample size in terms of number of households to be selected;

* $z$ is the statistic that defines the level of confidence desired; 

* $r$ is an estimate of a key indicator to be measured by the survey;

* $f$ is the sample design effect, deff, assumed to be 2.0 (default value); 

* $k$ is a multiplier to account for the anticipated rate of non-response; 

* $p$ is the proportion of the total population accounted for by the target population and upon which the parameter, r, is based; 

* $\hat{n}$ is the average household size (number of persons per household);

* $e$ is the margin of error to be attained.

$r$ is the most important variable for our purposes and at the present time it is uncertain what we should choose for our survey. For example we could choose $r=0.1$ if we believe we are trying to accurately estimate an attribute held by 10% of the population. For now we generate a range of possible values.

```{r}
r <- seq(0.1, 0.5, by=0.05)
r
```


Applying defualts recommended in the handbook for the other values gives:

```{r}
z <- 1.96
f <- 2
k <- 1.1
p <- 1
nhat <- 1
e <- 0.1*r
```

Our sample size estimates are then

```{r}
nh <- ((z^2)*r*(1-r)*f*k)/(p*nhat*(e^2))
as.data.frame(list("r"=r, "sample_size"=nh))
```

### Sample Size

For now, lets assume that we want to estimate the number of households in SAWPA with California-friendly landscaping, and let's further assume that between 30-35% of households in the region have this landscaping.

In that case, our sample size will need to be between `r ceiling(nh[5])` and `r ceiling(nh[6])`.

## Sample Generation

Generation of a sample of households for the Outdoor Water Efficiency Survey. The sample is generated using data stored primarily in the CaDC SCUBA data warehouse.

First connect to the database.
```{r}
library(DBI)
library(RPostgreSQL) 
source("../db_config.R")
db <- dbConnect(RPostgreSQL::PostgreSQL(), user=db_user, password=db_password, dbname=db_name, host=db_host, port=db_port)
# knitr::opts_chunk$set(connection = "db")
```

We want the ability to sample our tables, and the easiest wat to do that in Postgres is using the `TABLESAMPLE` command. `TABLESAMPLE` can only be used with actual tables and materialized views, so we will have to forgo our favorite [common table expressions](https://www.postgresql.org/docs/9.1/static/queries-with.html) (aka the `with` clause). Instead we create a new materialized view, or refresh it in this case as it has already been created.

```sql
CREATE MATERIALIZED VIEW sandbox.census_blockgroups_in_sawpa AS (
	SELECT c.*
	FROM census_blockgroup_polygons_2016 c, sandbox.sawpa_boundary_polygon s
	WHERE ST_Intersects(c.geom_3310, s.geom_3310)
)

--REFRESH MATERIALIZED VIEW sandbox.census_blocks_in_sawpa
```
```{sql, connection=db, tab.cap = "Number of census blockgroups in SAWPA"}
SELECT count(*) FROM sandbox.census_blockgroups_in_sawpa
```
Now we have a view of all the 2016 census blockgroups within the SAWPA service area. We can now sample from these to get the first level of our sample. A sample fraction of 1.5% gets us to approximately the number of block groups that we want. 

Next, the sampled block groups are joined with parcel polygons to get a list of all parcels within the sampled block groups. Additionally, these parcels are filtered to only parcels with an area less than 10,000 sqare meters (~2.5 acres). This does a decent, but not perfect job of filtering out large commercial and industrial parcels.

```sql
CREATE TABLE sandbox.parcels_in_blockgroup_sample AS 

WITH sawpa_blockgroup_sample AS (
  SELECT * 
  FROM sandbox.census_blockgroups_in_sawpa TABLESAMPLE BERNOULLI (1.5) REPEATABLE (1234)
),

parcels_in_blockgroup_sample AS (
  SELECT a.*, b.geoid as blockgroup_id
  FROM assessor_polygons a, sawpa_blockgroup_sample b
  WHERE ST_Intersects(a.geom, b.geom_3310)
  AND a.shape_area < 10000
)

SELECT * from parcels_in_blockgroup_sample
```
```{sql, connection=db, tab.cap = "Number of sampled blockgroups"}
SELECT count(*) 
FROM sandbox.census_blockgroups_in_sawpa TABLESAMPLE BERNOULLI (1.5) REPEATABLE (1234)
```
```{sql, connection=db, tab.cap = "Number of parcels in the sampled blockgroups"}
SELECT count(*) FROM sandbox.parcels_in_blockgroup_sample
```

<!-- ```{sql, connection=db, tab.cap = "Number of sampled census blockgroups"} -->
<!-- SELECT blockgroup_id, count(*) as parcel_count  from sandbox.parcels_in_blockgroup_sample -->
<!-- GROUP BY blockgroup_id -->
<!-- ORDER BY parcel_count -->
<!-- ``` -->

Finally, we sample from the parcels within the block groups to get the final household survey sample.
```sql
CREATE TABLE sandbox.hh_survey_sample AS
(SELECT * FROM sandbox.parcels_in_blockgroup_sample TABLESAMPLE BERNOULLI (10) REPEATABLE (1234))
```
```{sql, connection=db}
SELECT count(*) FROM sandbox.hh_survey_sample
```
